Critiquing the System Usability Scale from a Questionnaire Design Perspective: Beware of Acquiescence Bias

### 2013-09-16 paper_draft3.Rnw
created new shorter draft with all edits from gdoc

% Make sure hyperref comes last of your loaded packages, 
% to give it a fighting chance of not being over-written, 
% since its job is to redefine many LaTeX commands.
\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={Critiquing the System Usability Scale from a Questionnaire Design Perspective: Beware of Acquiescence Bias},
pdfauthor={XXXXXXXXX REPLACE WITH AUTHOR NAMES XXXXXXXXXX},
pdfkeywords={Usability, questionnaire, survey, System Usability Scale, SUS},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

@@

\numberofauthors{2}
\author{
    \alignauthor 1st Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Street}\\
    \affaddr{City}\\
    \email{E-mail address}
    \alignauthor 2nd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Street}\\
    \affaddr{City}\\
    \email{E-mail address}
}
% \author{
%   \alignauthor Ren\'{e} F. Kizilcec\\
%     \affaddr{Department of Communication}\\
%     \affaddr{Stanford University}\\
%     \email{kizilcec@stanford.edu}
%   \alignauthor Hendrik M\"{u}ller\\
%     \affaddr{User Experience Research}\\
%     \affaddr{Google, Inc.}\\
%     \email{hendrikm@google.com}
% }

@@

\begin{abstract}
The System Usability Scale (SUS) is probably the most widely employed measure of usability today. Numerous studies have assessed its psychometric properties and used it as a ``gold standard" in the development of alternative scales. Recent advances in questionnaire design research on satisficing, acquiescence, and other biases, however, now challenge some of the foundations of the SUS. In this note, we review literature on relevant survey biases, inspect each SUS item for such biases, and using a survey experiment, show that the SUS is vulnerable to significant acquiescence bias. We then propose a more robust scale, rooted in the SUS, which conforms with recent insights from questionnaire design research, and provide an example of how the proposed scale outperforms the SUS in terms of measurement sensitivity.
\end{abstract}

\keywords{Usability, questionnaire, survey, System Usability Scale, SUS}

\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}

@@

\subsubsection{Experiment setup}
Participants of a massive open online course offered by Stanford University were asked to complete an optional post-course survey. The survey received 1746 responses. 
%At the beginning of the survey respondents were asked to rate their overall experience with the course, their likelihood of taking another course with the same format, and the difficulty of the course. 
Respondents were randomly assigned to one of three weighted groups: 25\% were presented with the (original) SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received an example of a more robust scale proposed in this note (n=869) (see Table \ref{tab:items} for scale details). The system that respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos. The rest of the survey was the same for all respondents and contained typical course assessment questions.

@@

As the psychometric properties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics should be sufficient here. Table \ref{tab:dist} provides an overview of statistics and psychometric properties of the original and reversed SUS, and the more robust example scale. Although the SUS is frequently reported to have a higher coefficient $\alpha$ \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an $\alpha$ of 0.86 reflects a good level of internal consistency. Moreover, the factor analysis yields two eigenvalues greater than one, which is consistent with previous work on the SUS's factor structure \cite{lewis2009factor}.

\begin{table}[b]
\small
\centering
\begin{tabular}{llll}
\toprule
          & Original & Reversed & Alternative\\
Statistic & SUS      & SUS      & Scale$^\dagger$\\
\midrule
N     & 439 & 438 & 869\\
Range & [23, 100] & [8, 100] & [22, 100]\\
Mean  & 80.6 & 77.9 & 76.7\\
SD    & 16.1 & 14.2 & 14.7\\
Median& 85 & 80 & 78\\
IQR   & 20 & 18 & 22\\
Cronbach $\alpha$ & 0.86 & 0.73 & 0.67\\
$\left\vert{\lambda}>1\right\vert^{\ast}$ & 2 & 2 & 1\\
\bottomrule
\multicolumn{4}{l}{\scriptsize{$^{\ast}$number of eigenvalues greater than 1 in factor analysis}}\\
\multicolumn{4}{l}{\scriptsize{$^{\dagger}$proposed alternative items 3, 6, 9, 10 from Table \ref{tab:items}}}\\
\end{tabular}
\caption{Statistical and psychometric scale properties.}
\label{tab:dist}
\end{table}
<<eval=FALSE>>=
length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10])); alpha(m.orisus[,1:10])$total; fa(m.orisus[,1:10],1)$e.value
length(convSUS(m.revsus[,1:10])); range(convSUS(m.revsus[,1:10])); mean(convSUS(m.revsus[,1:10])); sd(convSUS(m.revsus[,1:10])); median(convSUS(m.revsus[,1:10])); IQR(convSUS(m.revsus[,1:10])); alpha(m.revsus[,1:10])$total; fa(m.revsus[,1:10],1)$e.value
length(convSUS2(m.sus2[,2:5])); range(convSUS2(m.sus2[,2:5])); mean(convSUS2(m.sus2[,2:5])); sd(convSUS2(m.sus2[,2:5])); median(convSUS2(m.sus2[,2:5])); IQR(convSUS2(m.sus2[,2:5])); alpha(m.sus2[,2:5])$total; fa(m.sus2[,2:5],1)$e.value
@

% We expect the  SUS to have high concurrent validity in the form of strong associations with related and weak associations with unrelated constructs. We find that the  SUS correlates weakly to moderately albeit significantly with the following related constructs: respondents' overall course experience ($r$=0.34, $t(437)$=8, $p\textless$0.001), and their likelihood to take another course with the same format ($r$=0.19, $t(437)$=4, $p\textless$0.001). However, there we find no significant association between the SUS and the perceived difficulty of the course ($r$=0.04, $t(437)$=0.9, p=0.4). [NEEDS MORE FRAMING. WHY IMPORTANT?]
% <<eval=FALSE>>=
% cor.test(m.orisus$oe,convSUS2(m.orisus[,1:10]))
% cor.test(m.orisus$redo,convSUS2(m.orisus[,1:10]))
% cor.test(m.orisus$learnsat,convSUS2(m.orisus[,1:10]))
% cor.test(m.orisus$difficulty,convSUS2(m.orisus[,1:10]))
% @
% 
% For the 439 responses to the SUS, Cronbach's $\alpha$ is 0.86 and the correlation of each item with the total score lies between 0.53 and 0.81 with 95\% confidence. Although the SUS is frequently reported to have a higher coefficient $\alpha$ \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an $\alpha$ of 0.86 reflects a good degree of interrelatedness.
% <<eval=FALSE>>=
% alpha(m.orisus[,1:10])$total
% omega(m.orisus[,1:10], nfactors=2)
% iclust(m.orisus[,1:10],nclusters=2)
% for(i in 1:10){
%     print(cor.test(convSUS2(m.orisus[,1:10]), m.orisus[,i])$conf[1:2])
% }
% @
% 
% A factor analysis of the 439 responses to the SUS suggests that the scale has a two-factor structure. A scree plot (Figure \ref{fig:scree}) illustrates that two factors have eigenvalues greater than one. This is consistent with previous work on the SUS factor structure \cite{lewis2009factor}.
% <<eval=FALSE>>=
% fa.parallel(m.orisus[,1:10])
% scree(m.orisus[,1:10])
% fa(m.orisus[,1:10],2)$e.values
% fa.diagram(fa(m.orisus[,1:10],2))
% @
% <<scree, fig.width=4, fig.height=3, fig.cap='Scree plot for the SUS and short SUS 2.0 showing that the SUS has a two-factor structure, while the short SUS 2.0 has a single-factor structure'>>=
% scree=data.frame(
%     n=c(1:10,1:4), 
%     eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,2:5],1)$e.values), 
%     sus=rep(c("Original SUS", "Short SUS 2.0"),c(10,4)))
% ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.65,.75))
% @

\subsubsection{Acquiescence Bias Conclusion}
A comparison between scores from the original and reversed SUS provides strong evidence that the SUS induces acquiescence bias. Without acquiescence bias, the average for each item on the original SUS would not be significantly different from the reverse-coded average for each item on the reversed SUS. However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in a significant difference between the original SUS average and reversed SUS reverse-coded average.

@@

\begin{table}[h]
\small
\centering
\begin{tabular}{rccccr}
\toprule
 & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Reversed} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
Item \# & M & SD & M & SD & p value \\
\midrule
1 & 3.09 & 0.90 & 3.21 & 1 & 0.002\\
2 & 3.28 & 1.06 & 3.23 & 0.86 & 0.006\\
3 & 3.30 & 0.84 & 3.45 & 0.89 & \textless0.001\\
4 & 3.54 & 0.92 & 3.08 & 1.23 & \textless0.001\\
5 & 3.00 & 0.92 & 3.10 & 1.01 & 0.013\\
6 & 3.26 & 1.02 & 2.89 & 1.14 & \textless0.001\\
7 & 3.12 & 0.90 & 3.09 & 0.95 & 0.868\\
8 & 3.10 & 1.18 & 3.33 & 0.84 & 0.173\\
9 & 3.29 & 0.89 & 3.45 & 0.91 & \textless0.001\\
10 & 3.26 & 1.06 & 2.35 & 1.49 & \textless0.001\\
\midrule
overall & 80.58 & 16.14 & 77.92 & 14.23 & \textless0.001\\
\bottomrule
\end{tabular}
\caption{Means, standard deviations, and p values from Mann-Whitney tests for each item and the overall score of the original and reversed SUS providing strong evidence that the SUS induces acquiescence bias.}
\label{tab:acqui}
\end{table}
<<eval=FALSE>>=
# testItems=function(i, oa, ra){
#     print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
# }
testItems=function(i, oa, ra){
    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(wilcox.test(oa,ra)$p.value,5), sep=" & "))
}
for(i in 1:10){
    testItems(i, m.orisus[,i], m.revsus[,i])
}
testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))

# subMean = function(data, indices) {mean(data[indices])}
# acDat=data.frame(item=rep(1:11, 2), sus=rep(c("Original SUS", "Reversed SUS"),each=11), est=NA, lwr=NA, upr=NA)
# acDat$kind = ifelse(acDat$item==11, "Overall Score", "Individual Item")
# 
# for(i in 1:10){
#     for(s in c("Original SUS", "Reversed SUS")){
#         if(s=="Original SUS"){
#             temp = boot(data=m.orisus[,i], statistic=subMean, R=10000)$t
#         } else {
#             temp = boot(data=m.revsus[,i], statistic=subMean, R=10000)$t
#         }
#      
#         acDat[acDat$item==i & acDat$sus==s, "est"] = mean(temp)
#         acDat[acDat$item==i & acDat$sus==s, "lwr"] = quantile(temp, 0.025)
#         acDat[acDat$item==i & acDat$sus==s, "upr"] = quantile(temp, 0.975)
#     }
# }
# temp = boot(data=convSUS(m.orisus[,1:10]), statistic=subMean, R=10000)
# ci = boot.ci(temp, type="bca")$bca[4:5]
# acDat[acDat$item==11 & acDat$sus=="Original SUS", "est"] = mean(temp$t)
# acDat[acDat$item==11 & acDat$sus=="Original SUS", "lwr"] = ci[1]
# acDat[acDat$item==11 & acDat$sus=="Original SUS", "upr"] = ci[2]
# 
# temp = boot(data=convSUS(m.revsus[,1:10]), statistic=subMean, R=10000)
# ci = boot.ci(temp, type="bca")$bca[4:5]
# acDat[acDat$item==11 & acDat$sus=="Reversed SUS", "est"] = mean(temp$t)
# acDat[acDat$item==11 & acDat$sus=="Reversed SUS", "lwr"] = ci[1]
# acDat[acDat$item==11 & acDat$sus=="Reversed SUS", "upr"] = ci[2]
# 
# acDat$item = ifelse(acDat$item==11, "all", acDat$item)
#     
# ggplot(acDat, aes(x=est, y=item, xmin=lwr, xmax=upr, linetype=sus)) +
#     geom_point(size=2) +
#     geom_errorbarh(height=.5) +
#     theme_bw() +
#     #scale_x_continuous(breaks=seq(0,100,2)) +
#     facet_wrap(~kind, scales="free", ncol=1) +
#     theme(legend.position="bottom", legend.key=element_blank(), strip.background=element_blank(), strip.text=element_text(face="bold"), legend.key.width=unit(1, "cm")) +
#     labs(x="Scores with 95% C.I.", y="", linetype="")

@

\section{Alternative Scale Proposal}

In light of the SUS's deficiencies uncovered in the previous section, we propose an updated set of items that, at their core, are equivalent to the SUS, but reduce vulnerability to survey biases, like acquiescence bias, satisficing, and hypothetical projections. The first step was to change the questionnaire items from being statements to questions in an effort to reduce acquiescence bias. Moreover, statements that were phrased as hypotheticals, such as item \#7 in Table \ref{tab:items}, were rephrased as concrete questions about the underlying construct referring back to the system that was just used by the respondent.

The second step was to change the scale from a Likert agreement scale to scales that reflect the relevant construct, such as ease/difficulty (item X), confidence (item X), learnability (item X), or complexity (item X). Following recommendations from question design research \cite{krosnickFabrigar1997}, unipolar constructs were presented as five-point scales, while bipolar constructs were presented as seven-point scales.

Our goal was to update the SUS items instead of creating a new measure, given that the SUS is probably the most established usability scale. These updated items with corresponding answer scales are presented in Table \ref{tab:items}. For our evaluation of the SUS 2.0, we were unable to use all ten items and opted for using a reduced number of items which we refer to as the short SUS 2.0. The short SUS 2.0 consists of four items marked with asterisks in Table \ref{tab:items} and covers the key dimensions of the SUS (confidence, ease of use, consistency, learnability).

The score calculation for the SUS 2.0 is different to that of the SUS, because the SUS 2.0 consists of six items with 5-point unipolar answer scales and four items with 7-point bipolar answer scales, instead of ten items on a 5-point scale. To calculate the SUS 2.0 score, first assign values between 0 -- 4 or 0 -- 6 depending on the number of scale points such that 0 reflects the worse usability response and 4 or 6 the best. Second, sum up the values for all ten responses to obtain an integer between 0 -- 48. Third, divide by 0.48 to obtain the SUS 2.0 score. For the short SUS 2.0, follow the same steps except that the sum of response values lies between 0 -- 18 and is divided by 0.18. In contrast to the SUS, non-responses are permitted, albeit not encouraged, in the SUS 2.0 and are accounted for by adjusting the denominator accordingly.

\subsection{Psychometric Properties of Alternative Scale}

@@

% \subsubsection{Concurrent Validity}
% Following the same procedure as for the SUS, we investigate the SUS 2.0's concurrent validity. We find that the short SUS 2.0 correlates moderately and significantly with respondents' overall course experience ($r$=0.31, $t(866)$=10, $p\textless$0.001) and their likelihood to take another course with the same format ($r$=0.21, $t(867)$=6, $p\textless$0.001); however, the short SUS 2.0 is only marginally associated with the perceived difficulty of the course ($r$=0.06, $t(867)$=1.7, p=0.08). [FRAMING!]
% <<eval=FALSE>>=
% cor.test(m.sus2$oe,rowSums(m.sus2[,2:5]))
% cor.test(m.sus2$redo,rowSums(m.sus2[,2:5]))
% cor.test(m.sus2$difficulty,rowSums(m.sus2[,2:5]))
% @
% 
% \subsubsection{Internal consistency}
% Cronbach's $\alpha$ for 869 responses to the short SUS 2.0 is 0.67 and individual item correlations with the total score vary between 0.54 and 0.84 with 95\% confidence. Coefficient $\alpha$ is smaller for the short SUS 2.0 than the SUS. However, given that coefficient $\alpha$ increases with the number of items, the ten-item SUS 2.0 is expected to have higher reliability than the four-item short SUS 2.0.
% <<eval=FALSE>>=
% alpha(m.sus2[,2:5])$total
% alpha(a.sus2[,2:10])
% scree(a.sus2[,2:10])
% fa.diagram(fa(a.sus2[,2:10],2))
% omega(m.orisus[,1:10],nfactors=2)
% for(i in 1:5){
%     print(cor.test(convSUS2(m.sus2[,2:5]), m.sus2[,i])$conf[1:2])
% }
% @
% 
% \subsubsection{Factor analysis}
% A factor analysis of short SUS 2.0 responses suggests that a single-factor structure is most appropriate for the scale. As illustrated in the scree plot (Figure \ref{fig:scree}), only one factor has an eigenvalues greater than one and the slope changes considerably at the two factor point. A scree test of SUS items that correspond to those in the short SUS 2.0 (items 1, 3, 7, 9) also yields a single-factor solution. Hence, the factor structure of the ten-item SUS 2.0 will likely resemble that of the SUS.
% <<eval=FALSE>>=
% scree(m.orisus[,c(1,3,5,7,9)])
% fa(m.sus2[,2:5],3)$loadings
% fa(m.sus2[,2:5],2)$loadings
% fa(m.sus2[,2:5],1)$loadings
% @

\subsection{Comparing Scale Sensitivity}
A good usability scale should exhibit a high level of sensitivity to reflect even subtle differences in usability. We conducted a second survey study in a post-course survey of an online course that ran on a different system (web interface) to investigate how an example of the proposed alternative scale (items 3, 6, 9, 10 from Table \ref{tab:items}) compares to the SUS in terms of sensitivity.

The two systems offered the same basic features, i.e. browsing and playing video lectures, but differed considerably in their design. We employed Molich and Nielsen's heuristic evaluation criteria \cite{molich1990improving} to informally establish which system has better usability. While both systems showed generally high usability, one system was deemed superior in four evaluation categories: match between system and the real world, consistency and standards, aesthetic and minimalist design, and help and documentation. This informal usability comparison was the basis for labeling one system as having `high usability' and the other `low usability'.

@@

%Figure \ref{fig:sens} illustrates usability ratings on the SUS and an example of the alternative scale for the high-usability and the low-usability system. As the usability scores from both scales were not normally distributed, 95\% confidence intervals were computed from 10,000 bootstrap replicates using the adjusted bootstrap percentile method.  While the SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the alternative scale exhibits good sensitivity. 
<<eval=FALSE>>=
# wilcox.test(convSUS(rbind(m.orisus[,1:10],m.revsus[,1:10])), convSUS(rbind(a.orisus[,1:10],a.revsus[,1:10])), conf.int=T)
# wilcox.test(rowSums(m.sus2[,2:5])/.18, rowSums(a.sus2[,2:5])/.18, conf.int=T)

wilcox.test(convSUS(m.orisus[,1:10]), convSUS(a.orisus[,1:10]), conf.int=T)
wilcox.test(sample(rowSums(m.sus2[,2:5])/.18,439), sample(rowSums(a.sus2[,2:5])/.18,96), conf.int=T)
@
% 
% <<sens, fig.height=2.5, fig.width=4.5, fig.cap='Evaluation of scale sensitivity for the SUS and alternative scale showing that both are sensitive enough to distinguish between a high and low usability system'>>=
% subMean = function(data, indices) {mean(data[indices])}
% m.oriBoot = boot(data=convSUS(rbind(m.orisus[,1:10],m.revsus[,1:10])), statistic=subMean, R=10000)
% a.oriBoot = boot(data=convSUS(rbind(a.orisus[,1:10],a.revsus[,1:10])), statistic=subMean, R=10000)
% m.sus2Boot = boot(data=rowSums(m.sus2[,2:5])/.18, statistic=subMean, R=10000)
% a.sus2Boot = boot(data=rowSums(a.sus2[,2:5])/.18, statistic=subMean, R=10000)
% 
% #  boot.ci(m.oriBoot, type="bca") # (78.21, 80.20 )  
% #  boot.ci(a.oriBoot, type="bca") # (80.96, 85.11 )  
% #  boot.ci(m.sus2Boot, type="bca") # (75.69, 77.63 )  
% #  boot.ci(a.sus2Boot, type="bca") # (78.78, 82.33 )  
% 
% sensDat=data.frame(
%     sus=c("SUS", "SUS", "Alternative\nScale", "Alternative\nScale"),
%     system=c("Low-Usability System","High-Usability System  ","Low-Usability System","High-Usability System  "),
%     score=c(m.oriBoot$t0, a.oriBoot$t0, m.sus2Boot$t0, a.sus2Boot$t0),
%     lwr=c(78.21, 80.96, 75.69, 78.78),
%     upr=c(80.20, 85.11, 77.63, 82.33))
%     
% ggplot(sensDat, aes(x=score, y=sus, xmin=lwr, xmax=upr, linetype=system)) +
%     geom_point(size=2) +
%     geom_errorbarh(height=.5) +
%     theme_bw() +
%     scale_x_continuous(breaks=seq(0,100,2)) +
%     theme(legend.position="bottom", legend.key=element_blank(), strip.background=element_blank(), strip.text=element_text(face="bold"), legend.key.width=unit(1, "cm")) +
%     labs(x="Scores with 95% C.I.", y="", linetype="")
% @

@@

This note's unique contribution is to critique the SUS from a questionnaire design perspective which has not been done before. We find strong evidence that the SUS induces acquiescence bias and show how a robust alternative scale with statements rephrased as questions and relevant answer scales achieves higher sensitivity in measuring usability than the SUS.

Future work should investigate the strength of association between the SUS and the proposed alternative scale, and establish the latter's reliability, validity, and factor structure. While more work on evaluating the proposed alternative scale is needed, the authors recommend using a more robust scale for measuring usability.

>>>>

% Make sure hyperref comes last of your loaded packages, 
% to give it a fighting chance of not being over-written, 
% since its job is to redefine many LaTeX commands.
\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={Critiquing the System Usability Scale from a Questionnaire Design Perspective: Beware of Acquiescence Bias},
pdfauthor={XXXXXXXXX REPLACE WITH AUTHOR NAMES XXXXXXXXXX},
pdfkeywords={System Usability Scale, SUS; response biases; usability evaluation; standardized questionnaires},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

@@

\numberofauthors{2}
\author{
    \alignauthor 1st Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Street, City}\\
    \email{E-mail address}
    \alignauthor 2nd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Street, City}\\
    \email{E-mail address}
}
% \author{
%   \alignauthor Ren\'{e} F. Kizilcec\\
%     \affaddr{Department of Communication}\\
%     \affaddr{Stanford University}\\
%     \email{kizilcec@stanford.edu}
%   \alignauthor Hendrik M\"{u}ller\\
%     \affaddr{User Experience Research}\\
%     \affaddr{Google, Inc.}\\
%     \email{hendrikm@google.com}
% }

@@

\begin{abstract}
The System Usability Scale (SUS) is undoubtedly the most widely employed measure of usability today. Numerous studies have assessed its psychometric properties and used it as a ``gold standard" in the development of alternative scale proposals. Recent advances in questionnaire design research on acquiescence and other biases, however, now challenge some of the foundations of the SUS. In this note, we review literature on relevant questionnaire design aspects, inspect the SUS for biases, and using a survey experiment, show that the SUS is vulnerable to significant acquiescence bias. We then propose an alternative scale, strongly rooted in the SUS, which conforms with recent insights from questionnaire design research, and provide an example of how the proposed scale outperforms the SUS in terms of measurement sensitivity.
\end{abstract}

\keywords{System Usability Scale, SUS; response biases; usability evaluation; standardized questionnaires}

%\terms{Human Factors; Design; Measurement}

\category{H.5.2.}{Information Interfaces and Presentation (e.g. HCI)}{User Interfaces–Evaluation/Methodology}

@@

\subsubsection{Experiment setup}
Participants of a massive open online course offered by Stanford University were asked to complete an optional post-course survey. The survey received 1746 responses. Respondents were randomly assigned to one of three weighted groups: 25\% were presented with the (original) SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received an example of a more robust scale proposed in this note (n=869) (see Table \ref{tab:items} for scale details). The system that respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos. The rest of the survey was the same for all respondents and contained typical course assessment questions.

@@

As the psychometric properties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics should be sufficient here. Table \ref{tab:dist} provides an overview of statistics and psychometric properties of the original and reversed SUS, and the alternative scale proposed below. A Cronbach's $\alpha$ of 0.86 reflects a good level of internal consistency, though the SUS is typically reported to have higher reliability \cite{bangor2008,lewis2009factor,borsci2009dimensionality}. A quick factor analysis yields two eigenvalues greater than one, which is consistent with previous work on the SUS's factor structure \cite{lewis2009factor}.

\begin{table}[b]
\small
\centering
\begin{tabular}{llll}
\toprule
          & Original & Reversed & Alternative\\
Statistic & SUS      & SUS      & Scale$^\dagger$\\
\midrule
N         & 439 & 438 & 869\\
Range     & [23, 100] & [8, 100] & [22, 100]\\
Mean (SD) & 80.6 (16.1) & 77.9 (14.2) & 76.7 (14.7)\\
Median (IQR) & 85 (20) & 80 (18) & 78 (22)\\
Cronbach $\alpha$ & 0.86 & 0.73 & 0.67\\
$\left\vert{\lambda}>1\right\vert^{\ast}$ & 2 & 2 & 1\\
\bottomrule
\multicolumn{4}{l}{\scriptsize{$^{\ast}$number of eigenvalues greater than 1 in factor analysis}}\\
\multicolumn{4}{l}{\scriptsize{$^{\dagger}$proposed alternative items 3, 6, 9, 10 from Table \ref{tab:items}}}\\
\end{tabular}
\caption{Statistical and psychometric scale properties.}
\label{tab:dist}
\end{table}
<<eval=FALSE>>=
length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10])); alpha(m.orisus[,1:10])$total; fa(m.orisus[,1:10],1)$e.value
length(convSUS(m.revsus[,1:10])); range(convSUS(m.revsus[,1:10])); mean(convSUS(m.revsus[,1:10])); sd(convSUS(m.revsus[,1:10])); median(convSUS(m.revsus[,1:10])); IQR(convSUS(m.revsus[,1:10])); alpha(m.revsus[,1:10])$total; fa(m.revsus[,1:10],1)$e.value
length(convSUS2(m.sus2[,2:5])); range(convSUS2(m.sus2[,2:5])); mean(convSUS2(m.sus2[,2:5])); sd(convSUS2(m.sus2[,2:5])); median(convSUS2(m.sus2[,2:5])); IQR(convSUS2(m.sus2[,2:5])); alpha(m.sus2[,2:5])$total; fa(m.sus2[,2:5],1)$e.value
@

\subsubsection{Acquiescence Bias Result}
A comparison between scores from the original and reversed SUS provides strong evidence that the SUS induces acquiescence bias. Without acquiescence bias, the average for each item on the original SUS would not be significantly different from the reverse-coded average for each item on the reversed SUS. However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in a significant difference between the original SUS average and reversed SUS reverse-coded average.

@@

\begin{table}[h]
\small
\centering
\begin{tabular}{rccccr}
\toprule
 & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Reversed} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
Item \# & M & SD & M & SD & p value \\
\midrule
1 & 3.09 & 0.90 & 3.21 & 1 & 0.002\\
2 & 3.28 & 1.06 & 3.23 & 0.86 & 0.006\\
3 & 3.30 & 0.84 & 3.45 & 0.89 & \textless0.001\\
4 & 3.54 & 0.92 & 3.08 & 1.23 & \textless0.001\\
5 & 3.00 & 0.92 & 3.10 & 1.01 & 0.013\\
6 & 3.26 & 1.02 & 2.89 & 1.14 & \textless0.001\\
7 & 3.12 & 0.90 & 3.09 & 0.95 & 0.868\\
8 & 3.10 & 1.18 & 3.33 & 0.84 & 0.173\\
9 & 3.29 & 0.89 & 3.45 & 0.91 & \textless0.001\\
10 & 3.26 & 1.06 & 2.35 & 1.49 & \textless0.001\\
\midrule
overall & 80.58 & 16.14 & 77.92 & 14.23 & \textless0.001\\
\bottomrule
\end{tabular}
\caption{Original and reversed SUS scores providing strong evidence for acquiescence bias.}
\label{tab:acqui}
\end{table}
<<eval=FALSE>>=
testItems=function(i, oa, ra){
    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(wilcox.test(oa,ra)$p.value,5), sep=" & "))
}
for(i in 1:10){
    testItems(i, m.orisus[,i], m.revsus[,i])
}
testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
@
\section{Alternative Proposal}

Based on the review and evaluation of the original SUS presented above, this section now discusses a proposal for an alternative scale. While reasoning based on recent questionnaire design research is used to change the wording of the different SUS items as well as the response scales, an experiment discussed its quality and sensitivity as compared to SUS.

\subsection{Proposed Wording Changes}

To minimize acquiescence bias in the SUS, each of the statements may be transformed into a construct-specific, neutral question with similarly neutral answer options matching the question construct. The appropriate scale and its length will then depend on the nature of the construct, i.e., if it is unipolar or bipolar in nature. To explain the reasoning that let to the proposed changes, this section now discusses a few exemplary items from the SUS. The intention is that the same consideration can then be applied to all other items, as summarized in Table \ref{tab:items}.

For example, to minimize acquiescence bias for item 9, we would first need to identify the underlying construct being asked about, which in this case is likely ``confidence". The leading statement can then easily be transformed into a construct-specific, neutral question, such as ``How confident were you using the system?". To determine the appropriate scale, it is critical to consider the polarity of ``confidence." As confidence naturally starts from a zero point (the absence of confidence), spans to a positive extreme (high confidence), without a natural negative extreme (i.e., several levels of negative confidence), this construct is unipolar in nature. As a result, it should then be measured on a 5-point, fully-labeled scale from ``Not at all confident" to ``Extremely confident" (see full scale in Table \ref{tab:items}).

As another example, item 3 refers to the bipolar construct of ease/difficulty. Bias may then be minimized by changing its wording to ``How easy or difficult was it to use the system?", giving ``easy" and ``difficult" equal weight so respondents are less led into either direction. Due to the bipolarity of this construct, a 7-point, fully labeled scale from ``Extremely difficult" to ``Extremely easy" may be used (see full scale in Table \ref{tab:items}). Note, as items 3 and 8 in the original SUS ask about the same underlying construct of ease/difficulty, they would result in the same reworded question, and hence, should only be included once in the full questionnaire.

Our goal was to propose alternative wording for the SUS items instead of creating a new measure, given that the SUS is probably the most established usability scale to date. Note that the purpose of this note is not an evaluation of the summarization of the questionnaire responses into a single score, hence, it is excluded from this discussion.

\subsection{Experimental Evaluation}

A good usability scale should exhibit a high level of sensitivity to reflect even subtle differences in usability. We conducted a second survey study in a post-course survey of an online course that ran on a different system (web interface) to investigate how an example of the proposed alternative scale (items 3, 6, 9, 10 from Table \ref{tab:items}) compares to the SUS in terms of sensitivity.

The two systems offered the same basic features, i.e. browsing and playing video lectures, but differed considerably in their design. We employed Molich and Nielsen's heuristic evaluation criteria \cite{molich1990improving} to informally establish which system has better usability. While both systems showed generally high usability, one system was deemed superior in four evaluation categories: match between system and the real world, consistency and standards, aesthetic and minimalist design, and help and documentation. This informal usability comparison was the basis for labeling one system as having ``high usability" and the other ``low usability".

\subsubsection{Psychometric Properties of the Alternative Scale}

@@

\subsubsection{Scale Sensitivity Result}

@@

<<eval=FALSE>>=
wilcox.test(convSUS(m.orisus[,1:10]), convSUS(a.orisus[,1:10]), conf.int=T)
wilcox.test(sample(rowSums(m.sus2[,2:5])/.18,439), sample(rowSums(a.sus2[,2:5])/.18,96), conf.int=T)
@

@@

This note's unique contribution is to critique the SUS from a questionnaire design perspective which has not been done before. We find strong evidence that the SUS induces acquiescence bias and show how an alternative scale with statements rephrased as questions and construct-specific, neutral answer scales achieve higher sensitivity in measuring usability than the SUS.

Future work could evaluate the SUS with respect to other biases, such as satisficing and the use of hypotheticals. Moreover, a thorough psychometric evaluation of the proposed alternative scale using a variety of systems is needed before we can recommend the adoption of this new scale. This will also require addressing the compatibility issue between old and new usability scores for the transition phase.
